{ \centering Scientific Report - Summary of Research\\ ~\\ Automatized compilation of sequential software to a diverse set of hardware accelerators\\ } ~\\ \vspace{18em} \large Grant number: PZ00P2 168016\\ Period: 01.05.2018– 31.12.2020\\ Grantee: Tobias Grosser\\ \today

### 2021-04-21 paper.tex
Report

\begin{enumerate}
\item Fast, robust, and correct polyhedral libraries
\item Solid polyhedral compiler infrastructure
\begin{itemize}
\item  Modular pattern-based optimizations for CPU and GPU systems
\item 
DeLICM: Scalar dependence removal
\end{itemize}
\item Fast GPU code generation demonstrated via a climate DSL
\item Code Generation for Emerging Architectures and FPGA
\end{enumerate}

@@

The central components of this library have already been upstreamed into the
open-source LLVM/MLIR project and are usable by the wider research and 
industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
\citet{pichanatan2021}.

@@

>>>>

\begin{enumerate}
\item Fast, robust, and correct polyhedral libraries
\item Automatic Raising to high-level descriptions
\item Fast GPU code generation demonstrated via a climate DSL
\item Code generation for emerging architectures and FPGA
\item Scalable performance analysis within a compiler.
\end{enumerate}

@@

The central components of this library have already been upstreamed into the
open-source LLVM/MLIR project and are usable by the wider research and 
industrial community. The work in this subproject has been published in \citet{grosser2020fast} and
is under review in \citet{pichanatan2021}.

\subsection{Automatic Raising to high-level descriptions}

Multi-level intermediate representations (IR) show great promise for lowering
the design costs for domain-specific compilers by providing a reusable,
extensible, and non-opinionated framework for expressing domain-specific and
high-level abstractions directly in the IR. But, while such frameworks support
the progressive lowering of high-level representations to low-level IR, they do
not raise in the opposite direction. Thus, the entry point into the compilation
pipeline defines the highest level of abstraction for all subsequent
transformations, limiting the set of applicable optimizations, in particular
for general-purpose languages that are not semantically rich enough to model
the required abstractions.  We propose Progressive Raising, a complementary
approach to the progressive lowering in multi-level IRs that raises from lower
to higher-level abstractions to leverage domain-specific transformations for
low-level representations. We further introduce MultiLevel Tactics, our
declarative approach for progressive raising, implemented on top of the MLIR
framework, and demonstrate the progressive raising from affine loop nests
specified in a general-purpose language to high-level linear algebra
operations.  Our raising paths leverage subsequent high-level domain-specific
transformations with significant performance improvements.

All this work has been has already been upstreamed into the open-source
LLVM/MLIR project and is usable by the wider research and industrial community.
The work in this subproject has been published in \citet{chelini2021mlir}.

@@

\subsection{Scalable performance analysis within a compiler}

Performance models are well-known instruments to understand the scaling
behavior of parallel applications. They express how performance changes as key
execution parameters, such as the number of processes or the size of the input
problem, vary. Besides reasoning about program behavior, such models can also
be automatically derived from performance data. This is called empirical
performance modeling. While this sounds simple at the first glance, this
approach faces several serious interrelated challenges, including expensive
performance measurements, inaccuracies inflicted by noisy benchmark data, and
overall complex experiment design, starting with the selection of the right
parameters. The more parameters one considers, the more experiments are needed
and the stronger the impact of noise. In this paper, we show how taint
analysis, a technique borrowed from the domain of computer security, can
substantially improve the modeling process, lowering its cost, improving model
quality, and help validate performance models and experimental setups.

In a subsequent step, we analyzed the optimality of function inlining as a means
to minimize code size and alter expose further optimization opportunities. This
work has shown promising code size reductions on SPEC 2017 and the LLVM compiler.
We hope to use it in the future as an enabling transformation for large-scale
automatic accelerator mapping.

The work in this subproject has been published in \citet{copic2021} and is under review in
\citet{theo2021}.
